Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence System

### 2020-02-14 Revisions/Rebuttal.md

---
title: "Rebuttal"
fontfamily: fourier
linestretch: 1
fontsize: 10pt
output:
  pdf_document:
    latex_engine: xelatex
---

<!--
rmarkdown::render('./Revisions/Rebuttal.md')
-->

<!-- =-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
# Letter to the editor
<!-- =-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

Dear Dr. Otto,

We are submitting a revised version of the manuscript "Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence System" (ID: 469411) for publication in Frontiers in Marine Science. We would like to thank the reviewer for their very constructive comments. We feel that we were able to address all of their comments and are now submitting an improved manuscript. Below we explain how we have revised our manuscript based on the reviews.

We thank you for considering our revised manuscript and I am looking forward to hearing from you.

Best regards,

David Beauchesne

<!-- =-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
# Question 1
<!-- =-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

***Please summarize the main findings of the study***

*This article presents 22 environmental drivers and uses these environmental drivers to identify areas of high cumulative exposure and characterize areas with similar cumulative exposure in the St. Lawrence System. In particular, the article introduces an open knowledge platform named eDrivers, though it doesn’t seem to be working at this moment (https://github.com/eDrivers - this site contains zero repositories, zero people, zero projects; the R package dDrivers is not available on CRAN). When this R package and the website become available, I believe the article and its related sources will be beneficial to the scientific and management communities. Despite its weaknesses (shown below), I believe this work is a good start for assessing and managing cumulative effects in the St. Lawrence System and can serve as a good example for other marine ecosystems.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
**Answer:**

> - The material has yet to be released publicly on GitHub as we were waiting for the review process to be completed. We can however provide access to the repositories in the interim.

> - As for the package, we will not be releasing a version on CRAN just yet. The material is not yet implemented on the SLGO servers (our partners on this project) and the data is currently contained directly in the package. Our goal is to publish the package and submit it for peer-review once the data is available and accessible through SLGO's API.
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

<!-- =-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
# Question 2
<!-- =-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

***Please highlight the limitations and strengths***

*Weaknesses (and potential solutions) are listed below:*

***Comment 1:***

*Authors need to make it clear up in the front that this work is only focusing on one of the four elements required to evaluate environmental impacts. Besides moving Lines 119-122 to around Line 77, it will be helpful to introduce the best practices in cumulative effects assessment and management: (1) selecting valued (ecosystem) components, (2) selecting indicators, (3) setting temporal and spatial scales, (4) addressing uncertainties, and (5) public participation. It will also be helpful to differentiate the two different indicators: effects-based and stressor-based. Effects-based indicators are direct measurements of the valued components (such as species abundance, biodiversity, etc.) and inherently encompass cumulative effects of multiple drivers. However, because of the poorly understood cause-effect linkages between stressors and valued components, it is currently hard to be used for decision-making. In comparison, stressor-based indicators (i.e., driver-based as this article defines and the type that this article uses) are usually better understood and can be proactively linked to management actions. However, these indicators do not capture the effects of all stressors and are not able to discern non-additive cumulative effects (e.g. synergism, antagonism). This point leads to the next weakness of this article.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Answer:***

> - We agree with reviewer 1 on the importance of clearly stating that we are solely focusing on drivers in our manuscript. The goal of our manuscript and the *eDrivers* initiative is to facilitate the widespread accessibility of such data for environmental assessments and integrative management. It is therefore clear to us that what we are presenting is but one element of a fully operational impact assessment. Our goal is however not to discuss cumulative effects assessments in particular. It is rather to provide knowledge on an element common to most holistic approaches, *i.e.* drivers characterization. We have strived to make this as clear as possible throughout the manuscript, namely in the introduction and in the perspectives sections, as noted by reviewer 1. As such, and considering the focus of the manuscript, we are hesitant to accept the suggestion of reviewer 1 to discuss cumulative effects assessment, best practices and indicator types at the beginning of the introduction. We feel that, considering the objective of this manuscript, the beginning of the introduction should rather focus on drivers specifically.

> - We nonetheless agree with reviewer 1's suggestions. We are therefore proposing the following modifications to the manuscript that we believe consolidates the essence of the suggestions made by reviewer 1:

  > 1. We modified the introduction to broadly state the major steps undertaken in holistic environmental approaches in general (lines 91 to 96).

  > 2. We modified the statement of the objectives of the manuscript at the end of the introduction to clearly state that we are focusing on drivers (lines 119 to 124).

  > 3. We modified the pre-existing paragraph presenting the importance of valued components in the perspective sections (lines 487 to 509). It is now presenting *eDrivers* as a suite of stressor-based indicators that hold certain advantages, but fail to capture effects on valued components and non-additive effects. We present effects-based indicators and their importance for understanding cumulative effects. We then argue that both are jointly required to properly diagnose causes of ecosystem change (Dube 2003, Environmental Impact Assessment Review 23(6):723-745; Jones 2016, Environmental Reviews 24(2):187-204).

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

***Comment 2:***

*Because of this inherent weakness of stressor-based indicators, the dealing and conclusion of this article related to driver interactions (Sections 3.4 and 4.2) are inadequate. How different stressors (or drivers) interact to create nonlinear, non-additive cumulative effects are dependent on a specific valued component of interest. This article simply assumes the summed intensity of two drivers as interactions, which is inaccurate and misleading. Consequently, the results based on the two drivers (hypoxia and fisheries) are not convincing (Section 4.2). How to create indicators that truly reflect negative impacts is essential and can be specific to a specific value component. For instance, temperature positive anomalies may have negative impact on certain species, it may not necessarily degrade ecosystem production. While the calculation of drivers intensity in Appendix is only one way of doing it, it is important for the computer package to provide users with flexibility to define their own indicators (i.e., one potential solution to the weakness).*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Answer:***

> - We begin by thanking the reviewer for pointing out this obvious weakness in the manuscript. We realize now that, even though we did not want to mislead readers in thinking we were evaluating driver interactions, we are in fact doing just that. As mentionned in reviewer 1 comments, driver co-occurrence provides no information on valued component-specific driver interactions and non-additive effects. Under the assumption that driver co-occurrence is necessary for driver interactions to exist, we believe it is reasonable to predict that driver interactions and non-additive effects will most likely be observed in areas where drivers overlap at high relative intensities. These areas indeed delineate where the effects of both drivers are most likely to impact valued components simultaneously. In presenting sections 3.4 and 4.2, our goal was to provide an illustration of how a simple two-drivers case could be used to identify areas of high relative intensity driver overlap. This should be used as segway to the more integrative cumulative footprint and hotspot indicators. As rightfully identified by reviewer 1, however, the manuscript is indeed misleading on the topic of driver interactions. We have therefore made the following modifications to the manuscript:

  > 1. We removed objective 2 from the manuscript and mentions of driver interactions, except in the context of discussing the likelihood on observing interactions in areas of driver co-occurrence.

  > 2. The methodology and results/discussions sections on driver interactions were moved to sections 3.3 and 4.1, respectively, and renamed *Driver co-occurrence*. We then modified the section as a simplified example of driver co-occurrence (lines XXX to XXX). We however did not modify the results/discussion section. The only mention of driver interactions was in the title itself. In the discussion, we merely discuss the areas of high intensity co-occurrence between the two drivers and present adverse effects that these drivers may have on species individually or simultaneously. We did however add a closing statement to the section stating that individual and interactive effects of both drivers is species- and community-dependent and that the co-occurrence of drivers in itself is not an evidence of driver interaction. We then state that assuming that driver co-occurrence is a prerequisite for driver interactions, however, it is reasonable to expect interactions to occur in areas where both drivers co-occur at high relative intensities (lines XXX to XXX)

> - Ultimately, this portion of the manuscript should serve as a simple illustration to better understand the complete cumulative footprint indicator. In this sense, if the reviewer still feels that proposed changes are not sufficient, we propose removing these sections from the manuscript.

> - As for providing flexibility in indicator definition, it is a feature that we wish to provide in future iterations of the package and the platform. We would however rather not include such enhancements at this point, as they are more technical in nature and should not preclude us from publishing this work. We however added a statement to the manuscript that presents this idea as a perspective and future work to be accomplished (lines XXX to XXX in perspectives).
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

***Comment 3:***

*Because of the paramount importance of defining valued components in the process of cumulative effects assessment and management, the absence of this element is the third weakness of this article. This weakness can potentially be overcome by adding “Future work”. In the future, similar work can be carried out to select and mapping valued components, which will allow explicit linkages between drivers and value components to be explored for better understanding cumulative effects.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Answer:***

> - We agree with reviewer 1 on the importance of valued components for cumulative effects assessments. Presenting a full cumulative impacts assessments and defining valued components is however not the objective of this manuscript, nor is it to discuss cumulative impacts in particular. The choices of valued components are fully dependent on the objective of a particular assessment or management scheme. In our experience, time and money is spent on recurring initiatives that seek to characterize sources of stress in a system for impacts assessments and integrative management. This is, in our view, unacceptable. Our goal was therefore to build an initiative that would limit this and provide access to formatted and comparable knowledge on sources of stress that are typically used for impact assessments and integrative management. This integrative dataset can then be used and adapted to the need of a user to assess impacts on a single or multiple valued components of interest, or the user can identify gaps in knowledge in the platform and ultimately contribute to the knowledge contained in the platform. We were therefore very deliberate in our decision not to include and discuss specific valued components. This manuscript is in fact part of a broader project that aims at evaluating cumulative impact on the ecological communities of the St. Lawrence. The description of drivers will thus be used on specific valued components in the St. Lawrence System.

> - As proposed by reviewer 1, we had already included a paragraph on the importance of valued components and their vulnerability to soures of stress in the conclusion. We however realize now that it was underwhelming in its ability to provide perspective and direction to how eDrivers may be used in the future. We therefore extended this paragragh, also in response to reviewer 1's first comment, to better discuss the significance of valued components and the different types of indicators for cumulative effects assessment (lines 487 to 509).
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

<!-- =-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->
# Question 3
<!-- =-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-= -->

***Please provide your detailed review report to the editor and authors (including any comments on the Q4 Check List):***

*This article is generally well written, presenting 22 environmental drivers. It uses these environmental drivers to identify areas of high cumulative exposure and characterize areas with similar cumulative exposure in the St. Lawrence System. In particular, the article introduces an open knowledge platform named eDrivers, though it doesn’t seem to be working at this moment (https://github.com/eDrivers - this site contains zero repositories, zero people, zero projects; the R package dDrivers is not available on CRAN). When this R package and the website become available, I believe the article and its related sources will be beneficial to the scientific and management communities. Despite its weaknesses (shown below), I believe this work is a good start for assessing and managing cumulative effects in the St. Lawrence System and can serve as a good example for other marine ecosystems.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Answer:***

> - See answer to question 1
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

*Weaknesses (and potential solutions) are listed below:*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Comment 1:***

*Authors need to make it clear up in the front that this work is only focusing on one of the four elements required to evaluate environmental impacts. Besides moving Lines 119-122 to around Line 77, it will be helpful to introduce the best practices in cumulative effects assessment and management: (1) selecting valued (ecosystem) components, (2) selecting indicators, (3) setting temporal and spatial scales, (4) addressing uncertainties, and (5) public participation. It will also be helpful to differentiate the two different indicators: effects-based and stressor-based. Effects-based indicators are direct measurements of the valued components (such as species abundance, biodiversity, etc.) and inherently encompass cumulative effects of multiple drivers. However, because of the poorly understood cause-effect linkages between stressors and valued components, it is currently hard to be used for decision-making. In comparison, stressor-based indicators (i.e., driver-based as this article defines and the type that this article uses) are usually better understood and can be proactively linked to management actions. However, these indicators do not capture the effects of all stressors and are not able to discern non-additive cumulative effects (e.g. synergism, antagonism). This point leads to the next weakness of this article.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Answer:***

> - See answer to question 2
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

***Comment 2:***

*Because of this inherent weakness of stressor-based indicators, the dealing and conclusion of this article related to driver interactions (Sections 3.4 and 4.2) are inadequate. How different stressors (or drivers) interact to create nonlinear, non-additive cumulative effects are dependent on a specific valued component of interest. This article simply assumes the summed intensity of two drivers as interactions, which is inaccurate and misleading. Consequently, the results based on the two drivers (hypoxia and fisheries) are not convincing (Section 4.2). How to create indicators that truly reflect negative impacts is essential and can be specific to a specific value component. For instance, temperature positive anomalies may have negative impact on certain species, it may not necessarily degrade ecosystem production. While the calculation of drivers intensity in Appendix is only one way of doing it, it is important for the computer package to provide users with flexibility to define their own indicators (i.e., one potential solution to the weakness).*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Answer:***

> - See answer to question 2
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

***Comment 3:***

*Because of the paramount importance of defining valued components in the process of cumulative effects assessment and management, the absence of this element is the third weakness of this article. This weakness can potentially be overcome by adding “Future work”. In the future, similar work can be carried out to select and mapping valued components, which will allow explicit linkages between drivers and value components to be explored for better understanding cumulative effects.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
***Answer:***

> - See answer to question 2
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

*Comment on the methods, results and data interpretation: The method on dealing with driver interactions is inadequate, and the resultant conclusions with regard to interactions are not convincing.*

***Answer:***

> - See answer to question 2

***Minor comments:***

*Line 229: add “on” before the analysis.*

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->
> - This was modified in the text (lines XXX to XXX).
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  -->

*Lines 327-328: fix language issue.*

### 2020-02-14 Revisions/Beauchesne-eDriversMS.Rmd
first complete draft of rebuttal

---
title: Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence System
author: David Beauchesne^1,2,\*^, Rémi M. Daigle^2^, Steve Vissault^3^, Dominique Gravel^3^, Andréane Bastien^4^, Simon Bélanger^5^, Pascal Bernatchez^5^, Marjolaine Blais ^6^, Hugo Bourdages^6^, Clément Chion^7^, Peter S. Galbraith^6^, Benjamin Halpern^8,9^, Camille Lavoie^2^, Christopher W. McKindsey^6^, Alfonso Mucci^10^, Simon Pineault^11^, Michel Starr^6^, Anne-Sophie Ste-Marie^4^, Philippe Archambault^2^
fontsize: 12pt
font-family: Times New Roman
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
  word_document:
    toc: false
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{listings}
   - \usepackage[svgnames]{xcolor}
   # - \usepackage{titling}
   # - \pretitle{\begin{flushleft}}
   # - \posttitle{\end{flushleft}}
   - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{bbm}
   - \usepackage{float}
   - \usepackage{ulem}
   - \definecolor{blue}{HTML}{20cad5}
relativeurls: true
bibliography: Beauchesne-eDriversMS.bib
csl: ../frontiers.csl
link-citations: yes
---

@@

The uncertainty associated with complex driver interactions must therefore be
taken into account when investigating environmental impacts [@cote2016], yet
most research on driver effects in marine environments remains overwhelmingly
focused on single driver assessments [@obrien2019]. Increasing exposure and the
experiences of past ecological tragedies such as the collapse of cod fisheries
[@frank2005; @dempsey2018] and the decline of the beluga and right whale
populations [@plourde2014] together urge the need to characterize the distribution,
intensity and overlap between drivers in the St. Lawrence System. This will
provide critical information on areas most exposed to cumulative drivers and on
the interaction potential of drivers in the St. Lawrence. It is also a necessary
step towards the holistic and integrated management of the St. Lawrence System.

Gathering environmental data for large scale, systematic initiatives can,
however, be a very challenging and time consuming – not to say painful – process.
On one hand, there is an overwhelming and expanding wealth of data available.
Such information overload may inhibit our ability to make decisions based on
scientific information, promote massive effort duplication, disproportionately
appropriate research funds to certain sectors, and obscure knowledge gaps amid
a sea of information [@eppler2004]. On the other hand, crucial data are lacking
and remain largely unavailable or inaccessible for a variety of reasons,
including proprietary rights, lack of organizational time, capacity and
training, and in some rare cases unwillingness to share, curtailing our ability
for appropriate decision-making.

@@

Our main objective in this study is to characterize the distribution and
intensity of drivers in the St. Lawrence. More specifically, we aim to: 1)
identify areas of high cumulative exposure, 2) identify drivers that are likely
to interact in the St. Lawrence and 3) characterize areas with similar
cumulative exposure regimes.
Here we focus solely on drivers, one of four elements required to evaluate
environmental impacts, *i.e.* drivers, ecosystem components of interest (*e.g.*
habitats or species) and the exposure and the vulnerability of the ecosystem
components to drivers. We achieve these objectives with the development of an
open knowledge platform named *eDrivers*. The platform was designed to
facilitate collaboration, real-time assessments of cumulative exposure and to
evolve with the addition of information and threats to the St-Lawrence ecosystems.

@@

@@

## Driver interactions

Interactions among drivers were investigated using the cumulative footprint
($F$) between pairs of drivers (objective 2). The intensity at which pairs of
drivers co-occur was evaluated using a two-dimensional kernel density estimate.
As there are 231 pairwise combinations between 22 drivers, we focus on a single
example using hypoxia and demersal destructive fisheries, two drivers known to
occur mainly in deeper areas of the St. Lawrence and, hence, an interaction
between the effects of the two drivers could be anticipated.

@@

In order to better capture potential interactions between drivers (objective 2)
and to identify areas with similar exposure regimes (objective 3),
we identify threat complexes using a clustering approach [*e.g.* see @bowler2019].
We use the term clusters in presenting the methods, but use threat complex when
discussing the results on cumulative exposure regimes.

@@

@@

While we cannot ascertain that high exposure areas are the most impacted,
we can safely predict that these are the areas where studying ecosystem state
will be the most complex due to the uncertainty associated with driver
interactions, an uncertainty bound to increase rapidly with the
number of interacting drivers [@cote2016].

## Driver interactions

Hypoxia is mainly distributed in the Laurentian, Anticosti and Esquiman Channels,
with the head of the Channels most exposed (Figure \ref{kernel}A).
Demersal destructive fisheries are located along the Laurentian Channel, the
heads of the Anticosti and Esquiman Channels and around the Magdalen Islands
(Figure \ref{kernel}B). By combining both drivers, we can observe that hypoxia
and demersal destructive fisheries overlap mostly at high relative intensity
(Figure \ref{kernel}D) in the vicinity of the Anticosti Gyre and the heads of
the Esquiman and Anticosti Channels (Figure \ref{kernel}C, Box 1).

@@

While the focus of our paper has been on the description of drivers in the
St. Lawrence, drivers are but one of the 4 elements required to properly
evaluate environmental impacts. This single knowledge node should be weaved with
other, comparable, nodes required for integrated management approaches such as
species distribution and marine protected areas. Ultimately, all of these
knowledge nodes could be weaved together through social-ecological metanetworks
[@dee2017].

Significant effort is still needed to bring our vision to fruition. Foremost
is to maintain our efforts to foster collaborations, develop platform
content and identify key knowledge gaps. A fair and efficient organizational
structure will be developed in order to manage *eDrivers* as a community and
appropriate funding must be secured to continue building this community and
ensure the long-term viability of the initiative, although the partnership with
SLGO partly addresses this issue.

@@

@@

\begin{figure}[H]
\centering
\includegraphics{../Figures/kernel.png}
\caption{Interaction between the intensity of hypoxia and demersal destructive fisheries in the St. Lawrence. An index of hypoxia ($A$) was created using bottom-water dissolved oxygen between 2013 and 2017 (Blais et al., 2018). Demersal destructive fisheries ($i.e.$ trawl and dredges) ($B$) intensity was evaluated from fisheries catch data collected between 2010 and 2015 used to measure annual area weighted total biomass ($kg$) in i $km^2$ grid cells (DFO, 2016b). See supplementary materials for more information on specific methodologies. Relative hypoxic stress and demersal destructive fisheries intensity was summed ($C$) to visualize their combined spatial distribution and intensity. Finally, individual density and the co-intensity of hypoxia and demersal destructive fisheries was investigated with a two-dimensional kernel analysis ($D$).}
\label{kernel}
\end{figure}

>>>>

---
title: Next Generation Planning - Structuring and Sharing Environmental Drivers Data for the St. Lawrence System
author: David Beauchesne^1,2,\*^, Rémi M. Daigle^2^, Steve Vissault^3^, Dominique Gravel^3^, Andréane Bastien^4^, Simon Bélanger^5^, Pascal Bernatchez^5^, Marjolaine Blais ^6^, Hugo Bourdages^6^, Clément Chion^7^, Peter S. Galbraith^6^, Benjamin Halpern^8,9^, Camille Lavoie^2^, Christopher W. McKindsey^6^, Alfonso Mucci^10^, Simon Pineault^11^, Michel Starr^6^, Anne-Sophie Ste-Marie^4^, Philippe Archambault^2^
fontsize: 12pt
font-family: Times New Roman
output:
  pdf_document:
    toc: false
    number_sections: true
    keep_tex: true
  word_document:
    toc: false
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{listings}
   - \usepackage[svgnames]{xcolor}
   # - \usepackage{titling}
   # - \pretitle{\begin{flushleft}}
   # - \posttitle{\end{flushleft}}
   - \usepackage{lscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{bbm}
   - \usepackage{float}
   - \usepackage{ulem}
   - \definecolor{blue}{HTML}{20cad5}
relativeurls: true
bibliography: Beauchesne-eDriversMS.bib
csl: frontiers.csl
link-citations: yes
---

@@

The uncertainty associated with complex driver interactions must therefore be taken into account when investigating environmental impacts [@cote2016], yet most research on driver effects in marine environments remains overwhelmingly focused on single driver assessments [@obrien2019]. Increasing exposure and the experiences of past ecological tragedies such as the collapse of cod fisheries [@frank2005; @dempsey2018] and the decline of the beluga and right whale populations [@plourde2014] together urge the need to characterize the distribution, intensity and overlap between drivers in the St. Lawrence System. This will provide critical information on areas most exposed to \textcolor{red}{\sout{cumulative}} drivers \textcolor{red}{\sout{and on the interaction potential of drivers}} in the St. Lawrence. It is also a necessary step towards the holistic and integrated management of the St. Lawrence System.

\textcolor{blue}{Gathering knowledge is a major component of any holistic approach, which typically involve the selection and description of valued ecosystem component (\textit{e.g.} habitats or species) and driver indicators, the assessment of the exposure and vulnerability of valued components to drivers, the selection of a proper spatio-temporal scale, site monitoring, and public and stakeholder participation (Dubé and Munkittrick 2001).} <!-- [@dube2001] jones2016 ???? Noble2013??? -->
Gathering environmental \textcolor{red}{\sout{data}} \textcolor{blue}{knowledge} for large scale, \textcolor{red}{\sout{systematic}} \textcolor{blue}{holistic} initiatives can, however, be a very challenging and time consuming – not to say painful – process. On one hand, there is an overwhelming and expanding wealth of data available. Such information overload may inhibit our ability to make decisions based on scientific information, promote massive effort duplication, disproportionately appropriate research funds to certain sectors, and obscure knowledge gaps amid a sea of information [@eppler2004]. On the other hand, crucial data are lacking and remain largely unavailable or inaccessible for a variety of reasons, including proprietary rights, lack of organizational time, capacity and training, and in some rare cases unwillingness to share, curtailing our ability for appropriate decision-making.

@@

Our main objective in this study is to characterize the distribution and
intensity of drivers in the St. Lawrence. More specifically, we aim to: 1)
identify areas of high cumulative exposure\textcolor{red}{\sout{, 2) identify drivers that are likely to interact in the St. Lawrence}} and \textcolor{red}{\sout{3)}} \textcolor{blue}{2)} characterize areas with similar cumulative exposure regimes.
\textcolor{red}{\sout{Here we focus solely on drivers, one of four elements required to evaluate environmental impacts, *i.e.* drivers, ecosystem components of interest (*e.g.* habitats or species) and the exposure and the vulnerability of the ecosystem components to drivers.}}
\textcolor{blue}{Here we focus solely on drivers, aiming at better understanding their distribution in the St. Lawrence and at facilitating the widespread availability of their characterization for holistic assessments and management approaches.}
We achieve these objectives with the development of an
open knowledge platform named *eDrivers*. The platform was designed to
facilitate collaboration, real-time assessments of cumulative exposure and to
evolve with the addition of information and threats to the St-Lawrence ecosystems.

@@

## \textcolor{red}{\sout{Driver interactions}} \textcolor{blue}{Driver co-occurrence}

\textcolor{red}{\sout{Interactions among drivers were investigated using the cumulative footprint ($F$) between pairs of drivers (objective 2). The intensity at which pairs of drivers co-occur was evaluated using a two-dimensional kernel density estimate. As there are 231 pairwise combinations between 22 drivers, we focus on a single example using hypoxia and demersal destructive fisheries, two drivers known to occur mainly in deeper areas of the St. Lawrence and, hence, an interaction between the effects of the two drivers could be anticipated.}}
\textcolor{blue}{As an introduction to indicators of cumulative exposure (objective 1), we begin by presenting a simplifed 2-driver example focusing on hypoxia and demersal destructive fisheries, two drivers known to occur mainly in deeper areas of the St. Lawrence. Driver co-ccurrence was evaluated spatially by summing the scaled intensity of drivers in each grid cell. The intensity at which pairs of drivers co-occur was evaluated using a two-dimensional kernel density estimate to investigate the relative intensity at which these two drivers typically co-occur.}

@@

@@

In order to \textcolor{red}{\sout{better capture potential interactions between drivers (objective 2) and to}}identify areas with similar exposure regimes (objective 3), we identify threat complexes using a clustering approach [*e.g.* see @bowler2019]. We use the term clusters in presenting the methods, but use threat complex when discussing the results on cumulative exposure regimes.

@@

## \textcolor{red}{\sout{Driver interactions}} \textcolor{blue}{Driver co-occurrence}

Hypoxia is mainly distributed in the Laurentian, Anticosti and Esquiman Channels, with the head of the Channels most exposed (Figure \ref{kernel}A). Demersal destructive fisheries are located along the Laurentian Channel, the heads of the Anticosti and Esquiman Channels and around the Magdalen Islands (Figure \ref{kernel}B). By combining both drivers, we can observe that hypoxia and demersal destructive fisheries co-occur mostly at high relative intensity (Figure \ref{kernel}D) in the vicinity of the Anticosti Gyre and the heads of the Esquiman and Anticosti Channels (Figure \ref{kernel}C, Box 1).

\textcolor{blue}{The individual or interactive effect of both drivers is thus species- and community-dependent and the co-occurrence of drivers in itself is not an evidence of driver interaction. Assuming that driver co-occurrence is a prerequisite for driver interactions, however, it is reasonable to expect interactions to occur in areas where both drivers co-occur at high relative intensities. }

@@

While we cannot ascertain that high exposure areas are the most impacted, we can safely predict that these are the areas where studying ecosystem state will be the most complex due to the uncertainty associated with driver \textcolor{red}{\sout{interactions}} \textcolor{blue}{co-occurrence}, an uncertainty bound to increase rapidly with the number of \textcolor{red}{\sout{interacting}} \textcolor{blue}{co-occurring} drivers [@cote2016].

@@

\textcolor{red}{\sout{While the focus of our paper has been on the description of drivers in the St. Lawrence, drivers are but one of the 4 elements required to properly evaluate environmental impacts. This single knowledge node should be weaved with other, comparable, nodes required for integrated management approaches such as species distribution and marine protected areas. Ultimately, all of these knowledge nodes could be weaved together through social-ecological metanetworks (Dee et al. 2017).}}

\textcolor{blue}{A caveat to our initiative is that it focuses on a single element required for fully operational impact assessments. Data that provide knowledge on the exposure of ecosystems to drivers are called stressor-based indicators (Dube and Munkittrick 2001; Dubé 2003; Jones 2016). These indicators efficently identify potential local impacts and can be proactively linked to decision-making, yet assume complete knowledge of drivers and fail to diagnose impacts on valued components or non-additive effects (Dube and Munkittrick 2001; Dubé 2003). In contrast, effect-based indicators are direct measurements of valued components (*e.g.* species abundance and biodiversity) and inherently capture the effects of multiple drivers (Dube and Munkittrick 2001; Dubé 2003). While effect-based indicators are considered superior to stressor-based indicators, they fail to ascribe observed effects to specific drivers (Dube and Munkittrick 2001; Dubé 2003). Stressor-based and effect-based indicators are therefore both required to diagnose causes of ecosystem change (Dube 2003; Jones 2016). As a knowledge node on stressor-based indicators, *eDrivers* should be weaved with other, comparable, nodes describing valued ecosystem components that can be linked to drivers and allow a better understanding of cumulative impacts. Ultimately, multiple knowledge nodes could be weaved together through social-ecological metanetworks analyses (Dee et al. 2016).} <!-- [@dube2001; @dube2003; @jones2016; @dee2017] -->

Significant effort is still needed to bring our vision to fruition. Foremost
is to maintain our efforts to foster collaborations, develop platform
content and identify key knowledge gaps. A fair and efficient organizational
structure will be developed in order to manage *eDrivers* as a community and
appropriate funding must be secured to continue building this community and
ensure the long-term viability of the initiative, although the partnership with
SLGO partly addresses this issue. \textcolor{blue}{We also wish to provide users with enhanced capabilities and flexibility in using the interactive tool and R package. This could include creating automatic reports and more flexibility for user-defined driver-based indicators.}

@@

\begin{figure}[H]
\centering
\includegraphics{../Figures/kernel.png}
\caption{\textcolor{red}{\sout{Interaction}} \textcolor{blue}{Co-occurrence} between \textcolor{red}{\sout{the intensity of }}hypoxia and demersal destructive fisheries in the St. Lawrence. An index of hypoxia ($A$) was created using bottom-water dissolved oxygen between 2013 and 2017 (Blais et al., 2018). Demersal destructive fisheries ($i.e.$ trawl and dredges) ($B$) intensity was evaluated from fisheries catch data collected between 2010 and 2015 used to measure annual area weighted total biomass ($kg$) in i $km^2$ grid cells (DFO, 2016b). See supplementary materials for more information on specific methodologies. Relative hypoxic stress and demersal destructive fisheries intensity was summed ($C$) to visualize their combined spatial distribution and intensity. Finally, individual density and the co-intensity of hypoxia and demersal destructive fisheries was investigated with a two-dimensional kernel analysis ($D$).}
\label{kernel}
\end{figure}

@@
